---
title: "Syllabus"
output:
  html_document:
    df_print: paged
---

## About the Course {#about-the-course .smith}

#### Instructor

-   Scott LaCombe (`slacombe@smith.edu`).Wright Hall 201. My hours will be held on Tuesdays 11-12, Wednesdays 10-11, and Thurdsays 4:15-5:15. Tuesdays and Wednesdays are by appointment slots ([sign up here](https://calendar.google.com/calendar/u/0/selfsched?sstoken=UUVsekFOdmJUYzVyfGRlZmF1bHR8MGY0NWUyMWJlMTk3M2ExMzc4YTdmMzM1NGI0MDY5Zjc)), and Thursdays are "group" office hours, meaning I have Wright 238 reserved and no appointment is required.

#### Description and Goals

The principal objective of this course is to introduce the underlying theory and the practical problems that are encountered in using linear statistical models in real life situations. The main goal of this course is to learn the steps of the model building process including, understanding the nature of the dataset and deciding on the best type of regression model, estimating the regression coefficients and presenting the model, regression diagnostics and remedial measures, prediction, and inference. Depending on the variables of interest, this course is mainly focus on three types of regression models, simple linear regression, multiple linear regression and logistic regression. The majority of the analysis will be done in R. Ultimately, the students will also use Quarto to prepare reports including all calculations, plots and interpretations.

Student Learning Outcomes: After successfully completing this course the students should be able to: - Use the exploratory data analytic techniques to suggest a model, fit the model and explain the theoretical properties and justifications for estimation and testing for the model.

-   Carry out regression diagnostics to check the model assumptions.
-   Remedial measures to achieve the model assumptions.
-   Use the model selection techniques to choose the best and simplest model.
-   Use logistic regression techniques to analyze data with binary response.
-   Use the statistical software such as R to perform the analysis mentioned above.
-   Prepare reports with Quarto.

**Prerequisite**: A college-level introductory statistics course (e.g. SDS/PSY 201, GOV 203, SOC 203, SDS 220, ECO 220), a score of 4 or 5 on the AP Statistics Exam, or permission of the instructor. Students need to be familiar with basic statistical concepts including graphical and numerical descriptive statistics, the normal and t-distributions, hypothesis tests, and confidence intervals.

**Co-requisite If this is your first SDS course you need to be co-enrolled in SDS 100.**

#### Readings

-   *STAT2: Modeling with Regression and ANOVA*, Second Edition by Ann R. Cannon, George W. Cobb, and Bradley A. Hartlaub. ISBN-13: 978-1319054076.\

This course uses one textbook. This book is required and is available at the bookstore. You may also choose to acquire the book elsewhere, but you NEED the textbook for this course (I suggest buying a used copy online). You will not need online access for the purposes of this course. Additional recommended readings, such as applications of material covered in class in the news, will be posted on the course website.

#### Class Structure

All class sessions will be a combination of lecture, small group work, and discussion. We will spend class time discussing your questions, looking at other examples, and doing activities. With this structure, it is VERY important that you do the assigned readings before coming to class. To this end, I will be starting many class sessions with questions about the readings to encourage you to keep up with the reading and to synthesize your thoughts before we begin class. Many activities will be designed to give you experience using statistical software to do the extensive computations most statistical analyses require. The computer is faster and more accurate than we are at doing arithmetic and graphs, but we have to know what arithmetic and which graphs will be useful.

All assignments will be submitted via Moodle. Finally, this course will use Slack to communicate including announcements, discussions about events, group project channels, and other uses as the semester progresses. I encourage you to message me on slack for course communication.

## Policies {#Policies .unnumbered}

#### Attendance

In keeping with Smith's core identity and mission as an in-person, residential college, SDS affirms College policy (as per the Provost and Dean of the College) that students will attend class in person. SDS courses will not provide options for remote attendance. Students who have been determined to require a remote attendance accommodation by the Office of Disability Services will be the only exceptions to this policy. As with any other kind of ADA accommodations, please notify your instructor during the first week of classes to discuss how we can meet your accommodations.

Participation has two components: 1) being present and 2) engaging in class activities. Your participation is an important part of the process of learning statistics, and we need you in class to help stimulate discussion. You can make a valuable contribution to the discussion by bringing up connections between our work and study designs you have seen in other courses, in the newspaper, or in research literature. If you miss class, I ask that you make up for your participation by posting something on slack such as a question about course material, article that relates to course content, or anything else related to course content. I will upload lecture slides on the course website, and I encourage you to get notes from another student and come to office hours with questions. You get one free absence that won't affect your participation grade (no questions asked and no need to contact me).

#### Collaboration

Much of this course will operate on a collaborative basis, and you are expected and encouraged to work together with a partner or in small groups to study, complete homework assignments, and prepare for exams. However, **every word that you write must be your own unless I'm explicitly asking you to submit a group submission**. Copying and pasting sentences, paragraphs, or blocks of code from another student is not acceptable and will receive no credit. No interaction with anyone but the instructor and course assistant is allowed on any exams. All students, staff and faculty are bound by the Smith College Honor Code.

#### Academic Honor Code Statement

Cases of dishonesty, plagiarism, or other infractions, will be reported to the Academic Honor Board. You must always provide appropriate citations for others' work and ideas. Giving other scholars due credit in your written and oral communication is a fundamental social gesture in academic work---a way for us to acknowledge each other's scholarship and signify that we respect each other.

From the [Smith honor code website](https://www.smith.edu/sao/handbook/socialconduct/honorcode.php):

> Smith College expects all students to be honest and committed to the principles of academic and intellectual integrity in their preparation and submission of course work and examinations. Students and faculty at Smith are part of an academic community defined by its commitment to scholarship, which depends on scrupulous and attentive acknowledgement of all sources of information, and honest and respectful use of college resources. **Cases of dishonesty, plagiarism, etc., will be reported to the Academic Honor Board.**

#### Classroom Environment

Realizing the benefits of a diverse space can only occur if we create a climate of psychological safety (Edmondson, 1999). To this end, we will always be respectful of one another. Together we should have the goal of creating an environment where we all feel comfortable sharing our thoughts and opinions. To this end, I value **"half-formed,"** informal thoughts---sometimes a deeper understanding is reached via communicating ideas before they are perfectly polished. If you have a question, I guarantee there is at least one other student with the same, so please speak up.

#### Accommodations

Everyone should have all that they need to succeed in this course. Please send me your accommodation letter, or have the Disability Office work with me. If you need to register for accommodations, please contact the Disability Services office at ODS\@smith.edu. Please check out the office [website](https://www.smith.edu/ods/) for more information. Smith provides flexible and individualized accommodation to students with documented disabilities that may affect their ability to fully participate in course activities or to meet course requirements. To receive accommodation services, students must be registered with the Smith College Disability Services office, College Hall 104; phone: (413) 585-2071 (voice, TTY, TDD). Students with disabilities are encouraged to contact the instructor for a confidential discussion of their individual needs for academic accommodation.

If you have special needs concerning quiz taking, please bring documentation of your needs and make an appointment to discuss them with me, at least ONE WEEK BEFORE the first exxam. That will give me time to provide accommodation for your needs.

## Assignments {#assignments .unnumbered}

1.  **Homework** \[20%\]: Homework assignments will be given during most weeks of the semester and will be problem sets from the textbook. Assignments will be posted on Thursdays after class and be due by the following Friday by 11:59 PM. Homework should be completed in Quarto; both the .QMD file and resulting html files should be submitted. Homework will be graded. For every day late on the homework assignments I will reduce your grade for that assignment by 10%. **I will drop the lowest homework grade at the end of the semester.**

2.  **Participation** \[5%\]: Active participation in class and regular attendance will comprise 5% of your grade. Most class periods will begin with questions from the reading for that day. Participating in the discussion about the readings is a great way to demonstrate engagement. Reading before class will help us move more quickly through content, and focus on the materials you found the most challenging. I will be tracking attendance through a sign up sheet each day, but there is no formal penalty for missed attendance. Attendance will help inform participation grades. As a reminder, participation includes in class work, but also posting on slack, coming to office hours, etc. There are many ways to fully participate in the course.

3.  **Exams** \[30%\]: There will be two exams throughout the semester using self-scheduled exams in McConnel Hall. Exams will be mostly short answer, with some asking you to solve a problem set, and others interpreting a regression table, as well as short essay answers explaining concepts. While the second exam is not explicitly cumulative, the content in this course builds on itself.

4.  **In-Class R Activities** \[5%\]: The R activities are assigned without any prior notice. The goal of these in-class activities is to give timely feedback on your understanding of the current material before the course moves on to new material. These projects are open book and sometimes be assigned as group work. However, each student is expected to write their own code and a report including the reasonable interpretation. Finalized activities are due at 11:59 PM the Sunday after they are assigned.

5.  **Case Studies** \[15%\]: There will be 4 case studies assigned this semester. These activities will give you a chance to *individually* grapple with a dataset, complete some sort of analysis, and interpret your results. A key component of these case studies is that you are correctly formatting your document with clear variable names, axes labels, hidden code, etc. This is a chance for you to practice creating documents in quarto/markdown. Secondly, I will be evaluating your for accurate and concise interpretation of your analysis.

6.  **Final Project and Participation** \[25%\]: For your final project you will be conducting a statistical investigation on a topic of your group's choosing. These projects will be done in groups of 3 (your choice of partner). You will use an existing dataset to answer a research question using statistical analysis (either multiple regression, or logistic regression). This project will include several check ins throughout the semester, and we will have 10-minute visual presentations the last week of class, with the final project being due during finals week. Throughout the semester we will have class time discussing how to choose a research question, finding data, etc. The goal of this project is to give you a firsthand look at how to engage in statistical analysis.

#### Summary

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
Assignment <- c("Homework", "In Class R Activities", "Exams", "Final Project", "Case Studies", "Participation", "Final Grade")
`Percent of Total` <- c("20%", "5%", "30%", "25%", "15%", "5%", "100%")
`Due Date` <- c("Fridays after assigned", "Daily", "Weekends-Self Scheduled", "See project guide", "Fridays after assigned", "")
kable(cbind(Assignment, `Percent of Total`, `Due Date`))
```

#### Grading

When I grade your written work, I am looking for problem solutions that are technically correct and reasoning that is clearly explained. Numerically correct answers, alone, are not sufficient on homework or exams. I value neatness and brief, clear, well-organized answers that explain your thinking. If the grader cannot read or follow your work, they cannot give you credit for it. The grader will check each homework submission for completeness and grade a subset of the exercises. Homework answer keys will be posted on Moodle after the due date.

#### Assignment Late Policy

For every day late on the written assignments I will reduce your grade for that assignment by ten percent. I will not accept any late final group research reports.

#### Final Grade Brackets

```{r, echo=FALSE}
Grade <- c("A", "A-", "B+", "B", "B-", "C+", "C", "C-", "D+", "D", "E")
Percent <- c("95-100%", "90-94%", "87-89%",	"83-86%",	"80-82%", "77-79%", "73-76%", "70-72%", "60-66%", "67-69%", "59% and below")
kable(cbind(Grade, Percent))
```

## Resources {#resources .unnumbered}

#### Course Website and Moodle

The [course website](https://lacombe129.github.io/SDS290/) will be regularly updated with lecture slides, project information, assignments, and other course resources. Assignments and grades will be submitted via Moodle. You should check both regularly.

#### Computing

The use of the `R` statistical computing environment with the [RStudio](http://rstudio.org) interface is thoroughly integrated into the course. You have two options for using `RStudio`:

-   The **server** version of `RStudio` on the [web](http://rstudio.smith.edu). The advantage of using the server version is that all of your work will be stored in the cloud, where it is automatically saved and backed up. This means that you can access your work from any computer with a web browser (Firefox is recommended) and an Internet connection.
-   A **desktop** version of `RStudio` installed on your machine. The downside to this approach is that your work is only stored locally, and you will have to manage your own installation. Nevertheless, I still recommend this option because you do not need wifi or network access to work, and you know it is on your computer. I find collaboration easier with this method too.

Note that you do not have to choose one or the other -- you may use both. However, it is important that you understand the distinction so that you can keep track of your work. Both `R` and `RStudio` are free and open-source, and are installed on most computer labs on campus. Please see the [Resources](resources.html) page for help with `R`.

#### Use of Technology during Class

As we will always be using our computers to attend this course, I hope it goes without saying that while the class is in session, you should resist the temptation to use your computer or cell phone for personal email, web browsing, social media, or any activity that's not related to the class.

#### Writing

Your ability to communicate results, which may be technical in nature, to your audience, which is likely to be non-technical, is critical to your success as a data analyst. The assignments in this class will place an emphasis on the clarity of your writing. I encourage you to use the [Jacobson Center](https://www.smith.edu/academics/jacobson-center) for assistance in writing.

The Statistical & Data Sciences Program is committed to ensuring that our students learn the skills necessary to become exemplary writers within our field. To that end, we have adopted a curricular model called Writing Enriched Curriculum, which has enabled us to articulate the writing skills we hope our graduates will acquire. These include:

-   The ability to adapt voice and expectations to the requirements of different genres (e.g. blog posts vs. analytical essays)
-   The ability to follow a writing process that includes brainstorming, outlining, initial drafting, peer review, editing, and revising
-   The ability to write with clarity and precision, even about issues of ambiguity and uncertainty
-   The ability to prioritize the important parts of a process and/or project to communicate
-   The ability to communicate a research question and how analysis will support that question
-   The ability to create impactful figures and tables
-   The ability to code with documentation and comments that are correctly indented, use a consistent style, and are human-readable.

Much of what you do in this course will support your understanding and development of these skills. If you have any questions about them, or would like more help in this work, please contact me and/or make an appointment to take your work to the Jacobson Center for Writing on their website. In particular, you may choose to bring your work to Peer Writing Tutors Elina Gordon-Halpern or Vivian Wei, both SDS majors who tutor for the Jacobson Center. Their schedules are below, but you may contact Sara Eddy (seddy\@smith.edu) for more information about their schedules or how to make an appointment.

#### Extra Help

There are [Statistics TAs](https://www.smith.edu/qlc/tutoring.html?colStats=open#PanelStats) available from 7:00-9:00pm on Sunday--Thursday evenings in Sabin-Reed 301 (except on Tuesdays where they are in . In addition, the Spinelli Center for Quantitative Learning (Seeley 207) supports students doing quantitative work across the curriculum. Your fellow students are also an excellent source for explanations, tips, etc.

#### Logistics

Please do not hesitate to reach out to me if you are feeling overwhelmed, lost, or need any help. I can connect you to university resources including counseling or other assistance.

On a similar note, please reach out before assignments or other deadlines are due if you need extensions, which I am happy to give out. I can accommodate delays if you stay in contact with me. If you have not contacted me all late assignments will have 10% deducted each day, and no assignment will be accepted more than 4 days late.

I encourage you to post questions and collaborate on slack as a first option before email. I will first defer to students to answer each other's questions and moderate as I have found explaining concepts can often be even more useful for learning than to hear it explained to you. I encourage you to use Slack to communicate with me as well over email, and my response time will be faster via Slack. Slack is also a great resource for you to work with other students, share notes, and collaborate in general (you can create separate groups for direct messages either with individual students, or with groups).

#### Other Resources

Note: As an employee of the college I have a mandate to report any incident of sexual or domestic violence that I hear or find out about to the Title IX Coordinator. If you want to make a confidential report, use the EthicsPoint service submission instructions [here](https://www.smith.edu/about-smith/reporting).

Smith is committed to providing support services and reasonable accommodations to all students with disabilities. To request an accommodation, please register with the Disability Services Office at the beginning of the semester. See this [link](https://www.smith.edu/about-smith/disability-services/register) for instructions and please do not hesitate to reach out if you have any questions, concerns, or suggestions for how I can facilitate your learning in this course.

#### Code of Conduct

As the instructor for this course, I am committed to making participation in this course a harassment-free experience for everyone, regardless of level of experience, gender, gender identity and expression, sexual orientation, disability, personal appearance, body size, race, ethnicity, age, or religion. Examples of unacceptable behavior by participants in this course include the use of sexual language or imagery, derogatory comments or personal attacks, trolling, public or private harassment, insults, or other unprofessional conduct.

As the instructor I have the right and responsibility to point out and stop behavior that is not aligned to this Code of Conduct. Participants who do not follow the Code of Conduct may be reprimanded for such behavior. Instances of abusive, harassing, or otherwise unacceptable behavior may be reported by contacting the instructor.

Everyone is expected to adhere to this Code of Conduct in all settings for this course: lectures, office hours, tutoring hours, and over Slack.

This Code of Conduct is available [here](https://www.smith.edu/about-smith/offices/student-affairs/student-handbook/code-student-conduct).

#### Tentative Schedule

Please refer to the complete [day-to-day schedule](https://docs.google.com/spreadsheets/d/148AKHY_3sfFzWpG17-OkK2TiI5QiCSd9whtTZPhfKt0/edit?usp=sharing) for more detailed information.
